A Survey of Current Exoskeletons\\ and Their Control Architectures and Algorithms\\ (Draft 3.2)

### 2015-09-30 survey.tex
updated conclusion

All the (strength augmentation) exoskeletons we reviewed can be categorized into two general control categories.  The exoskeleton either directly measures user intent (e.g., from EMG data or human-exoskeleton interaction forces), or it estimates intent from measurements at exoskeleton joints and applies a sensitivity amplification strategy to shadow the wearer.  The first strategy is difficult to implement using standard sensing modalities.  For instance, EMG data is notoriously noisy and presents both modeling and calibration challenges.  It would also be extremely difficult to keep these sensors in place to obtain accurate readings in dynamic environments or combat scenarios.  Similarly, many avoid the use of load cells for measurement of human-exoskeleton interface forces and the complicated sensor network required to accurately obtain interaction forces / torques (humans are compliant thus rigid attachment assumptions are typically inadequate).

While sensitivity amplification strategies circumvent these sensing issues by estimating human intent from exoskeleton joint readings, they cannot distinguish between user generated torques and those produced by the environment.  As such, they end up amplifying noise and the user must react to stabilize the system.  This is also unacceptable for highly dynamic exoskeletons operating in contact-rich environments.

We emphasize that user intent is critical, and current sensing challenges must be addressed in order to develop dynamic exoskeletons capable of serving in combat scenarios.  As a stop-gap, it may be possible to combine low information density EMG (thresholded on/off muscle activation signals) and human force/torque or angular measurements to filter user intent.  Such an approach may prove useful in a hybrid strategy, filtering external disturbance forces during sensitivity amplification control.  As a longer term solution, we highlight research in the area of MEMs and nano-fabrication, which has lead to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure / force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability. We recommend further study and the development of more advanced sensory systems to address the challenges inherent in dynamic exoskeleton control.

In addition to capturing user intent, it is important to develop accurate models of the human/exoskeleton system.  Feedback strategies, especially those based on force/torque control loops and series elastic actuation, are limited in bandwidth. 
%
It is important that an exoskeleton's control bandwidth stays well above (typically $2-3\times$) the human control bandwidth ($2-5\,$Hz), to track dynamic motion.  Several of exoskeletons controllers come close to this human bandwidth threshold under no-load conditions, and so can interfere in dynamic tasks.
%
Highly dynamic behaviors require predictive, feedforward models to reduce reliance on feedback and to allow a lower-impedance control response.  
Such feedforward, model-based controllers depend heavily on accurate exoskeleton / human dynamics and so modeling efforts must be emphasized.

% Finally, we note that 
Many of the common model-based control methods in this survey employ feedback linearization and inverse dynamics strategies for dynamic exoskeleton control.  
These methods (incorrectly) assume the exoskeleton is fully actuated and require users to provide certain torques at uncontrolled joints.  There are more advanced control schemes that extend to underactuated systems and may apply here (e.g., optimization-based controllers or virtual model control).  However, we note that there is a balance between feedback and feedforward control.  The better one is, the less reliant the system is on the other.
Improved feedback and direct sensing of user intent / human interaction forces therefore also offers to reduce sensitivity to high accuracy models and the fully actuated assumptions in inverse dynamics control.

>>>>

The (strength augmentation) exoskeletons we reviewed can be categorized into two general control categories.  The exoskeleton either directly measures user intent (e.g., from EMG data or direct human-exoskeleton interaction forces), or it estimates intent from measurements at the exoskeleton joints and applies a sensitivity amplification strategy to shadow the wearer.

Directly measuring user intent is difficult to implement using standard sensing modalities.  For instance, EMG data is notoriously noisy and presents both modeling and calibration challenges.  It would also be extremely difficult to keep these sensors in place to obtain accurate readings in dynamic environments or combat scenarios.  For similar reasons, many systems avoid the use of load cells or complicated sensor networks to accurately obtain operator-robot interaction forces / torques.

Sensitivity amplification strategies circumvent many of the direct operator-robot sensing issues by estimating human intent from measurements of the exoskeleton hardware. The issue with this policy is that it cannot generally distinguish between user generated torques and those produced by external sources.  As such, they end up amplifying not only the user's intent, but also external disturbances, requiring the operator to expend valuable energy to stabilize the system.  Additionally, even under the assumption that external forces can be minimized or mitigated by the user, enabling the suit to be highly responsive to perceived user intent implies that the hardware needs to be operating near the margin of stability.  This is not an optimal strategy for highly dynamic behaviors, as the suit may be pushed beyond the point to which the operator can reasonably stabilize it.

In light of these opinions, our main recommendation is that a control policy which implements a \emph{modified get out of the way} strategy would provide a reasonable strategy for dynamic exoskeleton control.  The get out of the way controller used in the Hydraulic Lower Extremity Exoskeleton discussed above would provide a starting point for this strategy.  This controller utilizes IMUs, foot contact sensors, and joint torque sensors to implement a user-reactive gravity compensating strategy.  Augmenting the sensor suit and control system to include sensors that directly measure the forces between the operator and the robot would enable this system not only avoid the user, but also to reject external disturbances.

To achieve this type of augmented get out of the way control, it may be possible to use low information density EMG (thresholded on/off muscle activation signals) with robot force/torque and angular measurements to filter operator-robot interaction forces.  As a longer term solution, we highlight research in the area of MEMs, nano-fabrication, soft sensors, etc., which has lead to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure / force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability.

We also highly stress the importance of high-bandwidth control loops.  It is important that an exoskeleton's control bandwidth stays well above (typically $2-3\times$) the human control bandwidth ($2-5\,$Hz), to track dynamic motion.  Several of exoskeletons controllers are close to lower limit of this human bandwidth threshold under no-load conditions, and so can interfere in dynamic tasks.  Note that these bandwidth issues are primarily a function of the exoskeleton hardware, and thus special care/analysis needs to be given/conducted to make sure that the hardware is capable of physically producing required behaviors.

Lastly, the importance of producing hardware that is physically capable of achieving dynamic performance is fundamentally linked to how mass is distributed throughout the system.  In general, this means that minimizing mass, more specifically distal masses on extremities, is desirable, which in turn means that at certain degrees of freedom must remain unactuated.  Underactuted control is a complicated problem which can either be ignored, in the sense that the control algorithm assumes the user will provide the necessary forces and torques for the unactuated degrees of freedom, dealt with using optimization-based control methods, or can be partially compensated using passive energy storage and dissipation elements.  Our recommendation is that a combination of biomechanically-inspired passive elements and optimization-based control methods will ultimately produce the best strategy for dealing with underactuation in higly-dynamic exoskeleton systems.

%Such feedforward, model-based controllers depend heavily on accurate exoskeleton / human dynamics and so modeling efforts must be emphasized.

% Finally, we note that 
%Many of the common model-based control methods in this survey employ feedback linearization and inverse dynamics strategies for dynamic exoskeleton control.  
%These methods (incorrectly) assume the exoskeleton is fully actuated and require users to provide certain torques at uncontrolled joints.  There are more advanced control schemes that extend to underactuated systems and may apply here (e.g., optimization-based controllers or virtual model control).  However, we note that there is a balance between feedback and feedforward control.  The better one is, the less reliant the system is on the other.
%
%Improved feedback and direct sensing of user intent / human interaction forces therefore also offers to reduce sensitivity to high accuracy models and the fully actuated assumptions in inverse dynamics control.
